Commutative Character Sheaves

### 2015-11-09 followup.tex
Commit for conversation with David

An $H$-equivariant linear character sheaf on $G$ is a triple $(\cs{L},\mu, \nu)$ where $(\cs{L},\mu)$ is a linear character sheaf and $\nu$ is compatible with $\mu$ in the sense. 
Consider the following morphisms: $c_0 : H\times G \times G\to H\times G \times H \times G$ defined by $c_0(h,g_1,g_2)\ceq(h,g_1,h,g_2)$;
$c_1 : H \times G\times G \to G\times G$ defined by $c_1(h, g_1, g_2) \ceq (a(h,g_1),g_2)$;  $c_2 : H \times G\times G \to H\times G$ defined by $c_2(h,g_1,g_2) \ceq (h, m(g_1,g_2))$; and $c_3 : H \times G\times G \to G\times G$ defined by $c_3(h, g_1, g_2) = (g_1, g_2)$. 
Then $(\cs{L},\mu,\nu)$ is an $H$-equivariant linear character sheaf on $G$ if $(\cs{L},\mu)$ is a linear character sheaf on $G$ and the following diagram of Weil local systems on $H \times G\times G$ commutes.
\[
\begin{tikzcd}
\arrow{d}[swap]{c_1^*(\mu)} c_1^* m^* \cs{L} & \arrow{l}[swap]{c_2\circ a = m\circ c_1}  c_2^* a^* \cs{L} \arrow{r}{c_2^*(\nu)} & c_2^* p^* \cs{L} \arrow{r}{p\circ c_2 = m\circ c_3}  & c_3^* m^*\cs{L} \arrow{d}{c_3^*(\mu)} \\
c_1^*(\cs{L}\boxtimes \cs{L}) \arrow{r}{p_i\circ c_1 = a\circ p_i\circ c_0} & c_0^*(a^*\cs{L}\boxtimes a^*\cs{L}) \arrow{r}{c_0^*(\nu\boxtimes \nu)} & c_0^*(p^*\cs{L} \boxtimes p^*\cs{L}) & \arrow{l}[swap]{p\circ p_i\circ c_0 = p_i \circ c_3}c_3^*(\cs{L}\boxtimes \cs{L})
\end{tikzcd}
\]
\todo{Check this.}
%In this diagram, the arrows not labeled are derived from canonical isomorphisms.
\iffalse
On stalks, this condition is
\[
\begin{tikzcd}
\arrow{d}[swap]{\mu_{(h\cdot g_1, h\cdot g_2)}} \cs{L}_{(h\cdot g_1)(h\cdot g_2)} & \arrow[equal]{l}  \cs{L}_{h\cdot(g_1 g_2)} \arrow{r}{\nu_{(h, g_2 g_2)}} & \cs{L}_{g_1 g_2} \arrow[equal]{r} & \cs{L}_{g_1 g_2} \arrow{d}{\mu_{(g_1,g_2)}} \\
\cs{L}_{h\cdot g_1}\otimes \cs{L}_{h\cdot g_2} \arrow[equal]{r} & \cs{L}_{h\cdot g_1}\otimes \cs{L}_{h\cdot g_2} \arrow{r}{\nu_{(h,g_1)}\otimes \nu_{(h,g_2)}} & \cs{L}_{g_1}\otimes \cs{L}_{g_2} & \arrow[equal]{l} \cs{L}_{g_1} \otimes \cs{L}_{g_2},
\end{tikzcd}
\]
for all $(h,g_1,g_2)\in H\times G\times G$.
\fi

@@

Now suppose $q : G \to G'$ is a regular epimorphism with kernel pair $(a,p)$ in the category of schemes
\[
\begin{tikzcd}
H\times G
 \arrow[shift left=2]{r}{a}
  \arrow[shift right=2,swap]{r}{p}
&
G 
\arrow{r}{q}
& 
G'
\end{tikzcd}
\]
Let $\cs{L}'$ be a linear character sheaf on $G'$.
Consider the functor
\[
q^* : \CS(G') \to \CS(G)
\]
given on objects by $(\cs{L}',\mu') \mapsto (q^*\cs{L}', (q\times q)^* \mu', \nu_{\cs{L}'})$; cf [Dictionary].
Then $q^*\cs{L}'$ is an $H$-equivariant linear character sheaf on $G$ when equipped with the canonical isomorphism
\[
\nu_{\cs{L}'} : a^* q^* \cs{L}' \mathop{\longrightarrow}\limits^{\iso} (q\circ a)^* \cs{L}' = (q\circ p)^*\cs{L}' \mathop{\longrightarrow}\limits^{\iso} p^* q^*\cs{L}.
\]
This simple observation shows that the functor $q^* : \CS(G') \to \CS(G)$  factors according to the following commuting diagram of functors
\[
\begin{tikzcd}
\CS(G) &&\arrow{ll}[swap]{q^*} 
\CS(G') \arrow{dl}{\text{`comparison'}}\\
& \arrow{ul}{\text{forget}} \CS_{H}(G). & 
\end{tikzcd}
\]

@@

\subsection{Forget}

Let us know specialize Section [] to the following case: $H = G\der$ and the action $a$ is the restriction of $m$ to $G\der\times G$.
Then $f : G\der \times G\to G\times G$ is simply the inclusion and $f_1: G\der\times G \to G$ is projection to the first coordinate followed by inclusion $G\der \hookrightarrow G$, so $f_1^*\cs{L} = \cs{L}\vert_{G\der} \boxtimes (\EE)_G$. 
Then $\beta : f_1^*\cs{L} \to (\EE)_{H\times G}$ determines an isomorphism $\beta : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$, and vice versa.

\todo{Under construction}

\begin{lemma}
The forgetful functor $\CS_H(G) \to \CS(G)$ is full and faithful with essential image $\CS'(G)$.\end{lemma}

\begin{proof}
\end{proof}

@@

As we have seen above, the fppf quotient $q : G\to G\ab = G/G\der$ exists in schemes.
Moreover, the quotient $q : G \to G\ab$ is faithfully flat and a $G\der$-torsor in the fppf topology, by\cite{SGA3}*{Expos\'e {VI}$_A$, Thm. 3.2}
%(In fact, it's a trivial torsor in that topology.)
It now follows from Lemma~[] that the comparison functor $\CS(G\ab)\to \CS_{G\der}(G)$ is an equivalence.
On the other hand, by Lemma~[], the forgetful functor $\CS_{G\der}(G)\to \CS(G)$ defines an equivalence $\CS_{G\der}(G)\to \CS'(G)$, where $\CS'(G)$ is the category of $(\cs{L},\mu,\nu)\in \CS_{G\der}(G)$ such that $\cs{L}\vert_{G\der} \iso (\EE)_{G\der}$.
It now follows that $q^* : \CS(G\ab) \to \CS(G)$ is fll and faithful with essential image $\CS'(G)$.
\end{proof}

@@

\begin{proof}[Proof of Theorem~\ref{thm:Gab}]
Combining Lemmas~[] and [] we see that $q^* : \CS(G\ab) \to \CS(G)$ is full and faithful.
% with essential image $\CS'(G)$, the category of linear character sheaves $\cs{L}$ on $G$ for which $\cs{L}\vert_{G\der} \iso (\EE)_{G\der}$. 
By Definition~[], this induces an equivalence between $\CCS(G\ab)$ and $\CCS(G)$. 
\end{proof}

>>>>

An $H$-equivariant linear character sheaf on $G$ is a triple $(\cs{L},\mu, \nu)$ where $(\cs{L},\mu)$ is a linear character sheaf and $\nu$ is compatible with $\mu$ in the sense. 
Consider the following morphisms: $c_0 : H\times G \times G\to H\times G \times H \times G$ defined by $c_0(h,g_1,g_2)\ceq(h,g_1,h,g_2)$;
$c_1 : H \times G\times G \to G\times G$ defined by $c_1(h, g_1, g_2) \ceq (a(h,g_1),a(h,g_2))$;  $c_2 : H \times G\times G \to H\times G$ defined by $c_2(h,g_1,g_2) \ceq (h, m(g_1,g_2))$; and $c_3 : H \times G\times G \to G\times G$ defined by $c_3(h, g_1, g_2) = (g_1, g_2)$. 
Then $(\cs{L},\mu,\nu)$ is an $H$-equivariant linear character sheaf on $G$ if $(\cs{L},\mu)$ is a linear character sheaf on $G$ and the following diagram of Weil local systems on $H \times G\times G$ commutes.
\[
\begin{tikzcd}
\arrow{d}[swap]{c_1^*(\mu)} c_1^* m^* \cs{L} & \arrow{l}[swap]{c_2\circ a = m\circ c_1}  c_2^* a^* \cs{L} \arrow{r}{c_2^*(\nu)} & c_2^* p^* \cs{L} \arrow{r}{p\circ c_2 = m\circ c_3}  & c_3^* m^*\cs{L} \arrow{d}{c_3^*(\mu)} \\
c_1^*(\cs{L}\boxtimes \cs{L}) \arrow{r}{p_i\circ c_1 = a\circ p_i\circ c_0} & c_0^*(a^*\cs{L}\boxtimes a^*\cs{L}) \arrow{r}{c_0^*(\nu\boxtimes \nu)} & c_0^*(p^*\cs{L} \boxtimes p^*\cs{L}) & \arrow{l}[swap]{p\circ p_i\circ c_0 = p_i \circ c_3}c_3^*(\cs{L}\boxtimes \cs{L})
\end{tikzcd}
\]
\iffalse
On stalks, this condition is
\[
\begin{tikzcd}
\arrow{d}[swap]{\mu_{(h\cdot g_1, h\cdot g_2)}} \cs{L}_{(h\cdot g_1)(h\cdot g_2)} & \arrow[equal]{l}  \cs{L}_{h\cdot(g_1 g_2)} \arrow{r}{\nu_{(h, g_2 g_2)}} & \cs{L}_{g_1 g_2} \arrow[equal]{r} & \cs{L}_{g_1 g_2} \arrow{d}{\mu_{(g_1,g_2)}} \\
\cs{L}_{h\cdot g_1}\otimes \cs{L}_{h\cdot g_2} \arrow[equal]{r} & \cs{L}_{h\cdot g_1}\otimes \cs{L}_{h\cdot g_2} \arrow{r}{\nu_{(h,g_1)}\otimes \nu_{(h,g_2)}} & \cs{L}_{g_1}\otimes \cs{L}_{g_2} & \arrow[equal]{l} \cs{L}_{g_1} \otimes \cs{L}_{g_2},
\end{tikzcd}
\]
for all $(h,g_1,g_2)\in H\times G\times G$.
\fi

@@

Now suppose $q : G \to G'$ is a regular epimorphism with kernel pair $(a,p)$ in the category of schemes
\[
\begin{tikzcd}
H\times G
 \arrow[shift left=2]{r}{a}
  \arrow[shift right=2,swap]{r}{p}
&
G 
\arrow{r}{q}
& 
G'
\end{tikzcd}
\]
Let $\cs{L}'$ be a linear character sheaf on $G'$.
Consider the functor
\[
q^* : \CS(G') \to \CS(G)
\]
given on objects by $(\cs{L}',\mu') \mapsto (q^*\cs{L}', (q\times q)^* \mu', \nu_{\cs{L}'})$; cf [Dictionary].
Then $q^*\cs{L}'$ is an $H$-equivariant linear character sheaf on $G$ when equipped with the canonical isomorphism
\[
\nu_{\cs{L}'} : a^* q^* \cs{L}' \mathop{\longrightarrow}\limits^{\iso} (q\circ a)^* \cs{L}' = (q\circ p)^*\cs{L}' \mathop{\longrightarrow}\limits^{\iso} p^* q^*\cs{L}.
\]
This simple observation shows that the functor $q^* : \CS(G') \to \CS(G)$  factors according to the following commuting diagram of functors
\[
\begin{tikzcd}
\CS(G) &\arrow{l}[swap]{q^*} 
\CS(G') \arrow{dl}{\text{`comparison'}}\\
\arrow{u}{\text{forget}} \CS_{H}(G). & 
\end{tikzcd}
\]

@@

\subsection{Quotient and restriction}

We now combine the material of the last two sections, by specializing to the following case: $H$ is a closed subgroup of $G$ and $q : G \to G/H$ is the fppf quotient.

\[
\begin{tikzcd}
\CS(H) & \arrow{l}[swap]{i^*} \CS(G) &\arrow{l}[swap]{q^*} 
\CS(G/H) \arrow{dl}{\text{`comparison'}} \\
 & \CS_{H}(G) \arrow{u}{\text{forget}}  & 
\end{tikzcd}
\]

\begin{lemma}
$\CS(G/H)$ is equivalent to the category of triples $(\cs{L},\mu,\beta)$ where $(\cs{L},\mu)\in \CS(G)$ and $\beta : \cs{L}\vert_H \to (\EE)_H$ is an isomorphism in $\CS(H)$.
\end{lemma}

In this case, the action $a$ is the restriction of $m$ to $H\times G$ and $f : H \times G\to G\times G$ is simply the inclusion and $f_1: H\times G \to G$ is projection to the first coordinate followed by inclusion $H \hookrightarrow G$, so $f_1^*\cs{L} = \cs{L}\vert_{H} \boxtimes (\EE)_G$. 
Then $\beta_1 : f_1^*\cs{L} \to (\EE)_{H\times G}$ determines an isomorphism $\beta : \cs{L}\vert_{H} \to (\EE)_{H}$, and vice versa.
Moreover, in this case, the constraint on $\beta$ is only this: $\beta : \cs{L}\vert_{H} \to (\EE)_{H}$ is an isomorphism in $\CS(H)$. 
%Let $\CS^H(G)$ denote the category of triples $(\cs{L},\mu,\beta)$ where $(\cs{L},\mu)\in \CS(G)$ and $\beta : \cs{L}\vert_H \to (\EE)_H$ is an isomorphism in $\CS(H)$.

@@

As we have seen above, the fppf quotient $q : G\to G\ab = G/G\der$ exists in schemes.
Moreover, the quotient $q : G \to G\ab$ is faithfully flat and a $G\der$-torsor in the fppf topology, by\cite{SGA3}*{Expos\'e {VI}$_A$, Thm. 3.2}
%(In fact, it's a trivial torsor in that topology.)
It now follows from Lemma~[] that the comparison functor $\CS(G\ab)\to \CS_{G\der}(G)$ is an equivalence.
On the other hand, by Lemma~[], the forgetful functor $\CS_{G\der}(G)\to \CS(G)$ defines an equivalence $\CS_{G\der}(G)\to \CS'(G)$, where $\CS'(G)$ is the category of $(\cs{L},\mu,\beta)$ where $(\cs{L},\mu)\in \CS(G)$ and $\beta : \cs{L}\vert_{G\der} \iso (\EE)_{G\der}$ in $\CS(G\der)$.
\end{proof}

@@

\begin{proof}[Proof of Theorem~\ref{thm:Gab}] 
\end{proof}
